---
title: "Nonparametric Analysis of US Dairy Production and Consumption"
subtitle: "GAM model"
author:
    - "Teo Bucci^[teo.bucci@mail.polimi.it]"
    - "Filippo Cipriani^[filippo.cipriani@mail.polimi.it]"
    - "Gabriele Corbo^[gabriele.corbo@mail.polimi.it]"
    - "Andrea Puricelli^[andrea3.puricelli@mail.polimi.it]"
output:
    pdf_document:
        toc: true
        toc_depth: 3
        number_section: true
        #keep_md: TRUE
    html_document:
        toc: true
        toc_float: true
        number_sections: true
date: "2023-02-17"
editor_options:
    chunk_output_type: inline
---
  
```{r setup, echo = FALSE}
knitr::opts_chunk$set(
    echo = TRUE,
    dev = c('pdf'),
    fig.align = 'center',
    fig.path = 'output/',
    fig.height = 6,
    fig.width = 12
)
```

# Load libraries and data

```{r ,message=FALSE}
library(pbapply)
library(mgcv)
library(conformalInference)
library(ggplot2)
library(progress)
library(parallel)
```

```{r echo = T, results = 'hide'}
data_path = file.path('data_updated_2021')
output_path = file.path('output')
data_infl <-read.table(
        file.path(data_path, 'production_facts_inflated.csv'),
        header = T,
        sep = ';'
    ) 
```

# Model

```{r}
model_gam = gam(
    avg_price_milk ~ s(dairy_ration, bs = 'cr')
    + milk_cow_cost_per_animal + milk_volume_to_buy_cow_in_lbs
    + milk_feed_price_ratio + s(milk_per_cow, bs = 'cr'),
    data = data_infl
)
```

```{r gam-smooth-terms}
par(mfrow = c(1,2))
plot(model_gam)
```

```{r}
milk_per_cow.grid=seq(range(data_infl$milk_per_cow)[1],
                      range(data_infl$milk_per_cow)[2],length.out = 100)
dairy_ration.grid=seq(range(data_infl$dairy_ration)[1],
                      range(data_infl$dairy_ration)[2],length.out = 100)
grid = expand.grid(
    milk_per_cow = milk_per_cow.grid,
    dairy_ration = dairy_ration.grid,
    milk_feed_price_ratio = mean(data_infl$milk_feed_price_ratio),
    milk_cow_cost_per_animal = mean(data_infl$milk_cow_cost_per_animal),
    milk_volume_to_buy_cow_in_lbs = mean(data_infl$milk_volume_to_buy_cow_in_lbs)
)
pred_gam = predict(model_gam, newdata = grid)
```

```{r gam-smooth-terms-surface}
plot3D::persp3D(
  x=milk_per_cow.grid,
  y=dairy_ration.grid,
  z=matrix(pred_gam, nrow=length(milk_per_cow.grid), ncol=length(dairy_ration.grid)),
  col.palette = heat.colors,
  #xlim = range(data_infl$milk_per_cow),
  xlab = 'milk_per_cow',
  ylab = 'dairy_ration',
  zlab = 'avg_price_milk',
  box = TRUE,
  #contour = TRUE,
  border='black',
  lwd=0.1,
  shade=0.1,
  bty="b2", # https://rdrr.io/cran/plot3D/man/perspbox.html
  phi = 20, theta = 50
)

with(
  data_infl,
  plot3D::points3D(
    milk_per_cow,
    dairy_ration,
    avg_price_milk,
    col = 'black',
    size = 1,
    pch=16,
    add=TRUE
  )
)
```

# Coefficients 

```{r}
tab = summary(model_gam)
format(as.data.frame(tab$p.coeff), scientific = FALSE)
as.data.frame(tab$s.table)
```

# Bootstrap interval on response

Taking into consideration the values of December, January and February of the covariates, we perform three bootstrap intervals on the prediction of the milk price, one for each month.

```{r}
milk_cow = c(1526.43,1531.21,1436.44)
dairy_rat = c(0.12308,0.12732,0.11571)
milk_feed = c(2.467,2.311,2.161)
milk_per_cow.med <- median(data_infl$milk_per_cow)
milk_volume_to_buy_cow_in_lbs.med <- median(data_infl$milk_volume_to_buy_cow_in_lbs)
```

```{r,echo=FALSE}
diagnostic_bootstrap = function(distro, obs){
  print(paste("Standard deviation: ", sd(distro)))
  print(paste("Bias: ", mean(distro) - obs))
  # computing confidence interval
  alpha <- 0.1
  right.quantile <- quantile(distro, 1 - alpha/2)
  left.quantile <- quantile(distro, alpha/2)
  # reverse-percentile
  CI <- c(obs - (right.quantile - obs),
          obs,
          obs - (left.quantile - obs))
  names(CI) <- c("lwr", "lvl", "upr")
  print(CI)
  plot(ecdf(distro), main='Parameter bootstrap distribution')
  abline(v = CI[2], lty=2)
  abline(v = CI[c(1,3)], lty=3)
  return(CI)
}

```

```{r}
CI <- matrix(0,3,3)
set.seed(1)
for(i in 1:3){
  newdata <-data.frame(milk_per_cow=milk_per_cow.med,
                       dairy_ration=dairy_rat[i],
                       milk_feed_price_ratio=milk_feed[i],
                       milk_cow_cost_per_animal=milk_cow[i],
                       milk_volume_to_buy_cow_in_lbs=milk_volume_to_buy_cow_in_lbs.med)
  B = 200
  fitted.obs <- predict(model_gam)
  res.obs <- data_infl$avg_price_milk - fitted.obs
  pred.obs = predict(model_gam, newdata = newdata)
  T.boot <- numeric(B)
  library(progress)
  pb <- progress_bar$new(
    format = "  processing [:bar] :percent eta: :eta",
    total = B, clear = FALSE)
  for (b in 1:B) {
    
    perm <- sample(1:nrow(data_infl), replace = T)
    dataset.boot = data_infl[perm,]
    
    model_gam_reduced.boot = 
      mgcv::gam(avg_price_milk ~ s(dairy_ration, bs = 'cr') 
                + milk_cow_cost_per_animal
                + milk_volume_to_buy_cow_in_lbs
                + milk_feed_price_ratio
                + s(milk_per_cow, bs = 'cr'), data = dataset.boot)
    
    T.boot[b] <- predict(model_gam_reduced.boot, newdata = newdata)
    pb$tick()
  }
  inter <- diagnostic_bootstrap(distro = T.boot, obs = pred.obs)
  CI[i,] <- inter
} 
```

and we compare them:

```{r,echo=FALSE}
jf=0.8
set.seed(1)
hist(data_infl$avg_price_milk,breaks=10,xlab='Milk Price',main = 'Prediction of milk prices')#,border=NA)
abline(v=jitter(CI[1,],jf),col=c('darkgray','black','darkgray'),lwd=c(2,3,2))
abline(v=CI[2,],col=c(2,'red',2),lwd=c(2,3,2))
abline(v=CI[3,],col=c('lightblue','blue','lightblue'),lwd=c(2,3,2))
```

```{r}
L = c(0.1790048 ,0.2006431 ,0.1819760 )
U = c(0.2928557 ,0.2928557 ,0.2516137 )
y = c(0.2424174 ,0.2425718 ,0.2157682 )
x=c("03 - Dec","02 - Jan","01 - Feb")
df = data.frame(x=x, y =y)

ggplot(df, aes(x = x, y = y)) +
    geom_errorbar(aes(ymax = U, ymin = L), width = 0.3) +
    geom_point(size = 4, col = "darkorange") +
    coord_flip() +
    labs(x = "Months",
         y = "Milk price [$/lbs]",
         title = "Prediction intervals") 
```


# Conformal Prediction

Using the \texttt{conformal.pred} function, it's possible to give a prediction and a conformal prediction interval on the price of the milk, considering fixed all variables except \texttt{milk\_per\_cow}.

```{r,echo=FALSE}
train_gam = function(x, y, out = NULL) {
  colnames(x) = c('var1', 'var2', 'var3', 'var4', 'var5')
  train_data = data.frame(y, x)
  model_gam = gam(y ~ s(var1, bs = 'cr') + s(var2, bs = 'cr') +
                    var3 + var4 + var5,data = train_data)
}
predict_gam = function(obj, new_x) {
  new_x = data.frame(new_x)
  colnames(new_x) = c('var1','var2','var3','var4','var5')
  predict.gam(obj, new_x)
}
```

The other 4 covariates are fixed to specified values.

```{r}
newdata <- c(milk_per_cow=0,dairy_ration=0.097,milk_feed_price_ratio=2.01,
             milk_cow_cost_per_animal=2037,milk_volume_to_buy_cow_in_lbs=10000)
milk_per_cow.grid=seq(range(data_infl$milk_per_cow)[1],
                      range(data_infl$milk_per_cow)[2],length.out = 100)
```

```{r}
wrapper_milk_per_cow=function(grid_point){
  newdata_t <- newdata
  newdata_t[1] <- grid_point
  alpha=0.1
  n_grid = 200
  c_preds = conformal.pred(
    cbind(
      data_infl$milk_per_cow,
      data_infl$dairy_ration,
      data_infl$milk_feed_price_ratio,
      data_infl$milk_cow_cost_per_animal,
      data_infl$milk_volume_to_buy_cow_in_lbs
    ),
    data_infl$avg_price_milk,
    newdata_t,
    alpha = alpha,
    verbose = T,
    train.fun = train_gam ,
    predict.fun = predict_gam,
    num.grid.pts = n_grid
  )
  inter<-c("LOWER" = c_preds$lo,
           "PRED" = c_preds$pred,
           "UPPER" = c_preds$up)
  return(inter)
}
```


```{r,eval=FALSE}
n_cores <- detectCores()
cl = makeCluster(n_cores)
invisible(clusterEvalQ(cl, library(DepthProc)))
clusterExport(cl, varlist = list("milk_per_cow.grid","wrapper_milk_per_cow",
                                 "newdata","data_infl","gam","predict.gam",
                                 "conformal.pred", "train_gam", "predict_gam"))
set.seed(1)
inter=pbsapply(milk_per_cow.grid,wrapper_milk_per_cow, cl = cl)
stopCluster(cl)
```

```{r, echo= FALSE}
load("ConfIntGam.RData")
```


```{r}
plot(milk_per_cow.grid,inter[2,],type='l',ylim=c(0.15,0.26))
points(milk_per_cow.grid,inter[1,],col=2,type='l')
points(milk_per_cow.grid,inter[3,],col=3,type='l')
```

