---
title: "Nonparametric Analysis of US Dairy Production and Consumption"
subtitle: "Permutation Tests for GAM"
author:
    - "Teo Bucci^[teo.bucci@mail.polimi.it]"
    - "Filippo Cipriani^[filippo.cipriani@mail.polimi.it]"
    - "Gabriele Corbo^[gabriele.corbo@mail.polimi.it]"
    - "Andrea Puricelli^[andrea3.puricelli@mail.polimi.it]"
output:
    pdf_document:
        toc: true
        toc_depth: 3
        number_section: true
        #keep_md: TRUE
    html_document:
        toc: true
        toc_float: true
        number_sections: true
date: "2023-02-17"
editor_options:
    chunk_output_type: inline
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
    echo = TRUE,
    dev = c('pdf'),
    fig.align = 'center',
    fig.path = 'output/',
    fig.height = 5,
    fig.width = 10
)
```

# Load libraries and data

```{r, message=FALSE}
library(pbapply)
library(mgcv)
```

```{r echo = T, results = 'hide'}
data_path = file.path('data_updated_2021')
output_path = file.path('output')
data_infl =
    read.table(
        file.path(data_path, 'production_facts_inflated.csv'),
        header = T,
        sep = ';'
    )
set.seed(1)
B = 1000
n = nrow(data_infl)
```

# $H_0$: `alfalfa_hay_price` $= 0$ VS $H_1$: `alfalfa_hay_price` $\neq 0$

```{r}
model_gam = gam(
    avg_price_milk ~ s(avg_milk_cow_number, bs = 'cr')
    + s(milk_per_cow, bs = 'cr')
    + s(dairy_ration, bs = 'cr')
    + milk_feed_price_ratio
    + milk_cow_cost_per_animal
    + milk_volume_to_buy_cow_in_lbs
    + alfalfa_hay_price
    + s(slaughter_cow_price, bs = 'cr'),
    data = data_infl
)
T0 = abs(summary(model_gam)$p.table[5, 3])
gam.H0 = gam(
    avg_price_milk ~ s(avg_milk_cow_number, bs = 'cr')
    + s(milk_per_cow, bs = 'cr')
    + s(dairy_ration, bs = 'cr')
    + milk_feed_price_ratio
    + milk_cow_cost_per_animal
    + milk_volume_to_buy_cow_in_lbs
    + s(slaughter_cow_price, bs = 'cr'),
    data = data_infl
)
res.H0 = gam.H0$residuals

wrapper = function() {
    permutation = sample(n)
    res.H0.perm = res.H0[permutation]
    Y.perm.H0 = gam.H0$fitted + res.H0.perm
    gam.perm = gam(
        Y.perm.H0 ~ s(avg_milk_cow_number, bs = 'cr')
        + s(milk_per_cow, bs = 'cr')
        + s(dairy_ration, bs = 'cr')
        + milk_feed_price_ratio
        + milk_cow_cost_per_animal
        + milk_volume_to_buy_cow_in_lbs
        + alfalfa_hay_price
        + s(slaughter_cow_price, bs = 'cr'),
        data = data_infl
    )
    return(abs(summary(gam.perm)$p.table[5, 3]))
}
T_H0 = pbreplicate(B, wrapper(), simplify = 'vector')
```


```{r perm-test-alfalfa_hay_price}
hist(sort(T_H0)[-1000],
     breaks = 100,
     main = 'Permutational distribution of test statistics',
     xlab = '')
abline(v = T0, col = 'red', lwd = 4)

plot(ecdf(sort(T_H0)[-1000]), main = 'ECDF of test statistics')
abline(v = T0, col = 'red', lwd = 4)
P = sum(T_H0 >= T0) / B
P
```

We cannot reject $H_0$ and proceed removing the covariate `alfalfa_hay_price` from the model.

# $H_0$: `avg_milk_cow_number` $= 0$ VS $H_1$: `avg_milk_cow_number` $\neq 0$

```{r}
model_gam = gam(
    avg_price_milk ~ s(avg_milk_cow_number, bs = 'cr')
    + s(milk_per_cow, bs = 'cr')
    + s(dairy_ration, bs = 'cr')
    + milk_feed_price_ratio
    + milk_cow_cost_per_animal
    + milk_volume_to_buy_cow_in_lbs
    + s(slaughter_cow_price, bs = 'cr'),
    data = data_infl
)
T0 = abs(summary(model_gam)$s.table[1, 3])
gam.H0 = gam(
    avg_price_milk ~ s(milk_per_cow, bs = 'cr')
    + s(dairy_ration, bs = 'cr')
    + milk_feed_price_ratio
    + milk_cow_cost_per_animal
    + milk_volume_to_buy_cow_in_lbs
    + s(slaughter_cow_price, bs = 'cr'),
    data = data_infl
)
res.H0 = gam.H0$residuals

wrapper = function() {
    permutation = sample(n)
    res.H0.perm = res.H0[permutation]
    Y.perm.H0 = gam.H0$fitted + res.H0.perm
    gam.perm = gam(
        Y.perm.H0 ~ s(avg_milk_cow_number, bs = 'cr')
        + s(milk_per_cow, bs = 'cr')
        + s(dairy_ration, bs = 'cr')
        + milk_feed_price_ratio
        + milk_cow_cost_per_animal
        + milk_volume_to_buy_cow_in_lbs
        + s(slaughter_cow_price, bs = 'cr'),
        data = data_infl
    )
    return(abs(summary(gam.perm)$s.table[1, 3]))
}
T_H0 = pbreplicate(B, wrapper(), simplify = 'vector')
```


```{r perm-test-avg_milk_cow_number}
hist(T_H0,
     breaks = 100,
     main = 'Permutational distribution of test statistics',
     xlab = '')
abline(v = T0, col = 'red', lwd = 4)

plot(ecdf(T_H0), main = 'ECDF of test statistics')
abline(v = T0, col = 'red', lwd = 4)
P = sum(T_H0 >= T0) / B
P
```

We cannot reject $H_0$ and proceed removing the covariate `avg_milk_cow_number` from the model.

# $H_0$: `slaughter_cow_price` $= 0$ VS $H_1$: `slaughter_cow_price` $\neq 0$

```{r}
model_gam = gam(
    avg_price_milk ~ s(milk_per_cow, bs = 'cr')
    + s(dairy_ration, bs = 'cr')
    + milk_feed_price_ratio
    + milk_cow_cost_per_animal
    + milk_volume_to_buy_cow_in_lbs
    + s(slaughter_cow_price, bs = 'cr'),
    data = data_infl
)
T0 = abs(summary(model_gam)$s.table[3, 3])
gam.H0 = gam(
    avg_price_milk ~ s(milk_per_cow, bs = 'cr')
    + s(dairy_ration, bs = 'cr')
    + milk_feed_price_ratio
    + milk_cow_cost_per_animal
    + milk_volume_to_buy_cow_in_lbs
    ,
    data = data_infl
)
res.H0 = gam.H0$residuals

wrapper = function() {
    permutation = sample(n)
    res.H0.perm = res.H0[permutation]
    Y.perm.H0 = gam.H0$fitted + res.H0.perm
    gam.perm = gam(
        Y.perm.H0 ~ s(milk_per_cow, bs = 'cr')
        + s(dairy_ration, bs = 'cr')
        + milk_feed_price_ratio
        + milk_cow_cost_per_animal
        + milk_volume_to_buy_cow_in_lbs
        + s(slaughter_cow_price, bs = 'cr'),
        data = data_infl
    )
    return(abs(summary(gam.perm)$s.table[3, 3]))
}
T_H0 = pbreplicate(B, wrapper(), simplify = 'vector')
```


```{r perm-test-slaughter_cow_price}
hist(T_H0,
     breaks = 100,
     main = 'Permutational distribution of test statistics',
     xlab = '')
abline(v = T0, col = 'red', lwd = 4)

plot(ecdf(T_H0), main = 'ECDF of test statistics')
abline(v = T0, col = 'red', lwd = 4)
P = sum(T_H0 >= T0) / B
P
```

We cannot reject $H_0$ and proceed removing the covariate `slaughter_cow_price` from the model.

# $H_0$: `milk_volume_to_buy_cow_in_lbs` $= 0$ VS $H_1$: `milk_volume_to_buy_cow_in_lbs` $\neq 0$

```{r}
model_gam = gam(
    avg_price_milk ~ s(milk_per_cow, bs = 'cr')
    + s(dairy_ration, bs = 'cr')
    + milk_feed_price_ratio
    + milk_cow_cost_per_animal
    + milk_volume_to_buy_cow_in_lbs,
    data = data_infl
)
T0 = abs(summary(model_gam)$p.table[4, 3])
gam.H0 = gam(
    avg_price_milk ~ s(milk_per_cow, bs = 'cr')
    + s(dairy_ration, bs = 'cr')
    + milk_feed_price_ratio
    + milk_cow_cost_per_animal,
    data = data_infl
)
res.H0 = gam.H0$residuals

wrapper = function() {
    permutation = sample(n)
    res.H0.perm = res.H0[permutation]
    Y.perm.H0 = gam.H0$fitted + res.H0.perm
    gam.perm = gam(
        Y.perm.H0 ~ s(milk_per_cow, bs = 'cr')
        + s(dairy_ration, bs = 'cr')
        + milk_feed_price_ratio
        + milk_cow_cost_per_animal
        + milk_volume_to_buy_cow_in_lbs,
        data = data_infl
    )
    return(abs(summary(gam.perm)$p.table[4, 3]))
}
T_H0 = pbreplicate(B, wrapper(), simplify = 'vector')
```


```{r perm-test-milk_volume_to_buy_cow_in_lbs}
hist(T_H0,
     breaks = 100,
     main = 'Permutational distribution of test statistics',
     xlab = '')
abline(v = T0, col = 'red', lwd = 4)

plot(ecdf(T_H0), main = 'ECDF of test statistics')
abline(v = T0, col = 'red', lwd = 4)
P = sum(T_H0 >= T0) / B
P
```

We reject $H_0$, maintaining the covariate `milk_volume_to_buy_cow_in_lbs`.

# $H_0$: `milk_cow_cost_per_animal` $= 0$ VS $H_1$: `milk_cow_cost_per_animal` $\neq 0$

```{r}
model_gam = gam(
    avg_price_milk ~ s(milk_per_cow, bs = 'cr')
    + s(dairy_ration, bs = 'cr')
    + milk_feed_price_ratio
    + milk_cow_cost_per_animal
    + milk_volume_to_buy_cow_in_lbs,
    data = data_infl
)
T0 = abs(summary(model_gam)$p.table[3, 3])
gam.H0 = gam(
    avg_price_milk ~ s(milk_per_cow, bs = 'cr')
    + s(dairy_ration, bs = 'cr')
    + milk_feed_price_ratio
    + milk_volume_to_buy_cow_in_lbs,
    data = data_infl
)
res.H0 = gam.H0$residuals

wrapper = function() {
    permutation = sample(n)
    res.H0.perm = res.H0[permutation]
    Y.perm.H0 = gam.H0$fitted + res.H0.perm
    gam.perm = gam(
        Y.perm.H0 ~ s(milk_per_cow, bs = 'cr')
        + s(dairy_ration, bs = 'cr')
        + milk_feed_price_ratio
        + milk_cow_cost_per_animal
        + milk_volume_to_buy_cow_in_lbs,
        data = data_infl
    )
    return(abs(summary(gam.perm)$p.table[3, 3]))
}
T_H0 = pbreplicate(B, wrapper(), simplify = 'vector')
```


```{r perm-test-milk_cow_cost_per_animal}
hist(T_H0,
     breaks = 100,
     main = 'Permutational distribution of test statistics',
     xlab = '')
abline(v = T0, col = 'red', lwd = 4)

plot(ecdf(T_H0), main = 'ECDF of test statistics')
abline(v = T0, col = 'red', lwd = 4)
P = sum(T_H0 >= T0) / B
P
```

We reject $H_0$, maintaining the covariate `milk_cow_cost_per_animal`.

# $H_0$: `milk_feed_price_ratio` $= 0$ VS $H_1$: `milk_feed_price_ratio` $\neq 0$

```{r}
model_gam = gam(
    avg_price_milk ~ s(milk_per_cow, bs = 'cr')
    + s(dairy_ration, bs = 'cr')
    + milk_feed_price_ratio
    + milk_cow_cost_per_animal
    + milk_volume_to_buy_cow_in_lbs,
    data = data_infl
)
T0 = abs(summary(model_gam)$p.table[2, 3])
gam.H0 = gam(
    avg_price_milk ~ s(milk_per_cow, bs = 'cr')
    + s(dairy_ration, bs = 'cr')
    + milk_cow_cost_per_animal
    + milk_volume_to_buy_cow_in_lbs,
    data = data_infl
)
res.H0 = gam.H0$residuals

wrapper = function() {
    permutation = sample(n)
    res.H0.perm = res.H0[permutation]
    Y.perm.H0 = gam.H0$fitted + res.H0.perm
    gam.perm = gam(
        Y.perm.H0 ~ s(milk_per_cow, bs = 'cr')
        + s(dairy_ration, bs = 'cr')
        + milk_feed_price_ratio
        + milk_cow_cost_per_animal
        + milk_volume_to_buy_cow_in_lbs,
        data = data_infl
    )
    return(abs(summary(gam.perm)$p.table[2, 3]))
}
T_H0 = pbreplicate(B, wrapper(), simplify = 'vector')
```


```{r perm-test-milk_feed_price_ratio}
hist(T_H0,
     breaks = 100,
     main = 'Permutational distribution of test statistics',
     xlab = '')
abline(v = T0, col = 'red', lwd = 4)

plot(ecdf(T_H0), main = 'ECDF of test statistics')
abline(v = T0, col = 'red', lwd = 4)
P = sum(T_H0 >= T0) / B
P
```

We reject $H_0$, maintaining the covariate `milk_feed_price_ratio`.

# $H_0$: `milk_per_cow` $= 0$ VS $H_1$: `milk_per_cow` $\neq 0$

```{r}
model_gam = gam(
    avg_price_milk ~ s(milk_per_cow, bs = 'cr')
    + s(dairy_ration, bs = 'cr')
    + milk_feed_price_ratio
    + milk_cow_cost_per_animal
    + milk_volume_to_buy_cow_in_lbs,
    data = data_infl
)
T0 = abs(summary(model_gam)$s.table[1, 3])
gam.H0 = gam(
    avg_price_milk ~ s(dairy_ration, bs = 'cr')
    + milk_feed_price_ratio
    + milk_cow_cost_per_animal
    + milk_volume_to_buy_cow_in_lbs,
    data = data_infl
)
res.H0 = gam.H0$residuals

wrapper = function() {
    permutation = sample(n)
    res.H0.perm = res.H0[permutation]
    Y.perm.H0 = gam.H0$fitted + res.H0.perm
    gam.perm = gam(
        Y.perm.H0 ~ s(milk_per_cow, bs = 'cr')
        + s(dairy_ration, bs = 'cr')
        + milk_feed_price_ratio
        + milk_cow_cost_per_animal
        + milk_volume_to_buy_cow_in_lbs,
        data = data_infl
    )
    return(abs(summary(gam.perm)$s.table[1, 3]))
}
T_H0 = pbreplicate(B, wrapper(), simplify = 'vector')
```


```{r perm-test-milk_per_cow}
hist(T_H0,
     breaks = 100,
     main = 'Permutational distribution of test statistics',
     xlab = '')
abline(v = T0, col = 'red', lwd = 4)

plot(ecdf(T_H0), main = 'ECDF of test statistics')
abline(v = T0, col = 'red', lwd = 4)
P = sum(T_H0 >= T0) / B
P
```

We reject $H_0$, maintaining the covariate `milk_per_cow`.

# $H_0$: `dairy_ration` $= 0$ VS $H_1$: `dairy_ration` $\neq 0$

```{r}
model_gam = gam(
    avg_price_milk ~ s(milk_per_cow, bs = 'cr')
    + s(dairy_ration, bs = 'cr')
    + milk_feed_price_ratio
    + milk_cow_cost_per_animal
    + milk_volume_to_buy_cow_in_lbs,
    data = data_infl
)
T0 = abs(summary(model_gam)$s.table[2, 3])
gam.H0 = gam(
    avg_price_milk ~ s(milk_per_cow, bs = 'cr')
    + milk_feed_price_ratio
    + milk_cow_cost_per_animal
    + milk_volume_to_buy_cow_in_lbs,
    data = data_infl
)
res.H0 = gam.H0$residuals

wrapper = function() {
    permutation = sample(n)
    res.H0.perm = res.H0[permutation]
    Y.perm.H0 = gam.H0$fitted + res.H0.perm
    gam.perm = gam(
        Y.perm.H0 ~ s(milk_per_cow, bs = 'cr')
        + s(dairy_ration, bs = 'cr')
        + milk_feed_price_ratio
        + milk_cow_cost_per_animal
        + milk_volume_to_buy_cow_in_lbs,
        data = data_infl
    )
    return(abs(summary(gam.perm)$s.table[2, 3]))
}
T_H0 = pbreplicate(B, wrapper(), simplify = 'vector')
```


```{r perm-test-dairy_ration}
hist(T_H0,
     breaks = 100,
     main = 'Permutational distribution of test statistics',
     xlab = '')
abline(v = T0, col = 'red', lwd = 4)

plot(ecdf(T_H0), main = 'ECDF of test statistics')
abline(v = T0, col = 'red', lwd = 4)
P = sum(T_H0 >= T0) / B
P
```

We reject $H_0$, maintaining the covariate `dairy_ration`.
