---
title: "Nonparametric Analysis of US Dairy Production and Consumption"
subtitle: "Functional Depth on clusters"
author:
    - "Teo Bucci^[teo.bucci@mail.polimi.it]"
    - "Filippo Cipriani^[filippo.cipriani@mail.polimi.it]"
    - "Gabriele Corbo^[gabriele.corbo@mail.polimi.it]"
    - "Andrea Puricelli^[andrea3.puricelli@mail.polimi.it]"
output:
    pdf_document:
        toc: true
        toc_depth: 3
        number_section: true
        #keep_md: TRUE
    html_document:
        toc: true
        toc_float: true
        number_sections: true
date: "2023-02-17"
editor_options:
    chunk_output_type: inline
---
  
```{r setup, echo = FALSE}
knitr::opts_chunk$set(
    echo = TRUE,
    dev = c('pdf'),
    fig.align = 'center',
    fig.path = 'output/',
    fig.height = 6,
    fig.width = 12
)
```

# Load libraries and data

```{r, message=FALSE}
library(BNPTSclust)
library(roahd)
library(fda.usc)
```

```{r echo = T, results = 'hide'}
data_path = file.path('data_updated_2021')
output_path = file.path('output')
dairy = read.table(file.path(data_path, 'dairy.csv'),header = T,sep = ';') 
```

Remove total cheese consumptions:

```{r}
dairy <- dairy[,-c(12,13,14,15)]
```

# Bayesian nonparametric clustering

```{r functional-consumption-exploratory}
years = 1975:2021
matplot(
    years,
    dairy,
    type = 'l',
    lty = 1,
    col = "black",
    main = "Consumption of different types of cheese",
    xlab = "Years",
    ylab = "Dairy consumption [lbs/person, normalized]"
)
```

Perform the clustering using the function \texttt{tseriescm} from the package \texttt{BNPTSclust}, based on a nonparamteric bayesian approach.

```{r, eval=FALSE, results = 'hide', message=FALSE}
tseriescm.out <-
    tseriescm(
        dairy,
        maxiter = 100,
        burnin = 10,
        thinning = 2,
        level = FALSE,
        trend = TRUE,
        seasonality = FALSE,
        priorb = TRUE,
        b = 0
    )
```

```{r, include=FALSE}
load("cluster.RData")
```

The clustering obtains five clusters, three of which non singular:

```{r functional-bayesian-clustering}
par(mfrow = c(1, 3))
matplot(
    years,
    dairy[, tseriescm.out$gnstar == 1],
    type = 'l',
    col = "black",
    xlab = 'Years',
    ylab = 'Dairy consumption [lbs/person, normalized]',
    main = "First Cluster"
)
matplot(
    years,
    dairy[, tseriescm.out$gnstar == 2],
    type = 'l',
    col = "black",
    xlab = 'Years',
    ylab = 'Dairy consumption [lbs/person, normalized]',
    main = 'Second Cluster'
)
matplot(
    years,
    dairy[, tseriescm.out$gnstar == 4],
    type = 'l',
    col = "black",
    xlab = 'Years',
    ylab = 'Dairy consumption [lbs/person, normalized]',
    main = 'Third Cluster'
)
```

# Functional Depth measures on clusters

```{r}
c1 = dairy[, tseriescm.out$gnstar == 1]
c2 = dairy[, tseriescm.out$gnstar == 2]
c3 = dairy[, tseriescm.out$gnstar == 4]
grid = 1:dim(dairy)[1]
c1 = as.matrix(c1)
c1_f = fData(grid, t(c1))
c2 = as.matrix(c2)
c2_f = fData(grid, t(c2))
c3 = as.matrix(c3)
c3_f = fData(grid, t(c3))
```

Compute the modified band depth

```{r}
median_curve1 <- median_fData(fData = c1_f, type = "MBD") 
median_curve2 <- median_fData(fData = c2_f, type = "MBD") 
median_curve3 <- median_fData(fData = c3_f, type = "MBD") 
```

```{r functional-bayesian-clustering-depth-median}
par(mfrow = c(1, 3))

plot(c1_f, col = "grey", xlab = "Years", ylab = "Dairy Consumption", main = "First Cluster")
grid_ecg1 <- seq(median_curve1$t0, median_curve1$tP, by = median_curve1$h)
lines(grid_ecg1, median_curve1$values, col = "blue", lwd = 2)

plot(c2_f, col = "grey", xlab = "Years", ylab = "Dairy Consumption", main = "Second Cluster")
grid_ecg2 <- seq(median_curve2$t0, median_curve2$tP, by = median_curve2$h)
lines(grid_ecg2, median_curve2$values, col = "red", lwd = 2)

plot(c3_f, col = "grey", xlab = "Years", ylab = "Dairy Consumption", main = "Third Cluster")
grid_ecg3 <- seq(median_curve3$t0, median_curve3$tP, by = median_curve3$h)
lines(grid_ecg3, median_curve3$values, col = "black", lwd = 2)
```
 We can note an up tren ecc.....

```{r eval=F, echo=T, results = 'hide', message=FALSE}
par(mfrow=c(1,3))
invisible(fbplot(c1_f))
invisible(fbplot(c2_f))
invisible(fbplot(c3_f))
```

There are no outliers in the 3 clusters.

```{r functional-outliergram, warning=FALSE}
invisible(outliergram(c1_f))
invisible(outliergram(c2_f))
invisible(outliergram(c3_f))
```

# Cluster analysis

```{r}
c1_names <- colnames(c1)
knitr::kable(c1_names, col.names="Cluster 1")
#knitr::kable(c1_names, col.names="Cluster 1", format = "latex")
c2_names <- colnames(c2)
knitr::kable(c2_names, col.names="Cluster 2")
#knitr::kable(c2_names, col.names="Cluster 2", format = "latex")
c3_names <- colnames(c3)
knitr::kable(c3_names, col.names="Cluster 3")
#knitr::kable(c3_names, col.names="Cluster 3", format = "latex")
```

